### 2021-06-10 abstractESSLLI21/abstractESSLLI1.Rmd
Adds abstract review

A considerable amount of literature exists on  bias detection and mitigation in NLP models, especially word2vec embeddings, which represent words as number vectors (see e.g. @bolukbasi2016man and @Gonen2019lipstick and references therein). The most common method used compares  cosine similarity between words from protected groups and attributes that are considered to be stereotypical or harmful in some way, and this method will be in our focus.

@@

Such methods are statistically problematic. One issue, specific to @Manzini2019blackToCriminal, is that no distinction is made based on whether given a protected word and a  class of attributes is stereotypically associated with this  protected word or a different protected word. A related, general problem, is that all the authors ignore the step of comparing their results with control groups, especially control groups of stereotype-neutral human attributes. That such a move is important is suggested for instance by Figure 1 prepared using the original word list for religion-related protected words extended with control attributes, where such human attributes are also closer to protected classes than 1.

Another, also serious problem, is that  all the authors calculate means of means and their authors run statistical tests on sets of means. Unforutnately, by pre-averaging the data they throw away information about sample sizes, and  they remove variation, and so pre-averaging  tends to manufacture false confidence.

@@

\noindent
\begin{tabular}{ll}
\begin{minipage}[c]{0.35\linewidth}
\scalebox{0.65}{
\begin{tabular}{cccccc}\toprule
$X$ to $A$  & $X$ to $B$ & $Y$ to $A$ & $Y$ to $B$ & $\sigma$  & WEAT \\
\midrule
0 & 0 & 0 & 0 & 0.05 & 1.82 \\
0 & 0 & 0 & 0 & 0.001 & -1.93\\
.1 & 0 & 0 & .1 & 0.05 & 1.49 \\
.1 & 0 & 0 & .1 & .1 & 1.22\\
\bottomrule
\end{tabular}
}
\end{minipage}& \begin{minipage}[c]{0.6\linewidth}
Two protected classes, $X=\{t_1,t_2\}$ and $Y=\{t_3,t_4\}$, and two five-element attribute sets $A$ and $B$. One simple simulation draws normally distributed values for such combinations for two cases in which the underlying mean similarities are in fact equal, two cases in which they are different, and the underlying variances are different (code available upon request).
\end{minipage}
\end{tabular}

@@

The following observations are worthwhile. (1) For points randomly drawn from distributions where there is no difference in means the calculated effect size can easily be  1.82, whereas the largest effect size reported by @Caliskan2017semanticsBiases is 1.81. (2) For samples from distributions where the means are different, the (absolute) effect sizes can easily be lower than in the first two simulations.  As Figure 2. illustrates,  quite some uncertainty is involved, far from what systematically low mean-based p-values reported in the papers might suggest. Part of the problem is random variation unaccounted for in the original approach (see Figures 3 and 4 for an example), and part of the problem is that  that  non-negligible changes in effect size can result from a shift in the standard deviation of the original process, because with the decrease of standard deviation the numerator in (WEAT) decreases.

@@

We build analogous models for other topic groups (race, gender), and different embeddings (Google, and hard-debiased Reddit). The general conclusion is that once we proceed this way, the situation is much less obvious. The word list sizes are small, sample sizes are small, and so posterior density intervals are wide. Some stronger bias can be discerned in the gender case. Moreover, sometimes the difference between associated, different and human, are not very impressive. The cost of debiasing turns out sometimes to be that   neutral human predicates get closer to protected classes (in religion),  change in gender after debiasing is really minor with   zero still out out of HPDI range, and debiasing improvement for race is a small improvement   achieved at the price of moving neutral terms closer to protected words (there is no space to visualise these results in this abstract).

>>>>

\todo{czy konieczne jest pisanie "number vectors"? jeżeli tak, to w licencjacie musze pozmieniac, bo tam czasami jest samo vectors, a czasami "they are assigned numerical values"}
A considerable amount of literature exists on  bias detection and mitigation in NLP models, especially word2vec embeddings, which represent words as number vectors (see e.g. @bolukbasi2016man and @Gonen2019lipstick and references therein). The most common method used compares  cosine similarity between words from protected groups and attributes that are considered to be stereotypical or harmful in some way, and this method will be in our focus.

@@

Such methods are statistically problematic. One issue, specific to @Manzini2019blackToCriminal, is that no distinction is made based on whether given a protected word and a  class of attributes is stereotypically associated with this  protected word or a different protected word. A related, general problem, is that all the authors ignore the step of comparing their results with control groups, especially control groups of stereotype-neutral human attributes. \todo{nie wiem, czy nie przeredagowałabym tego zdania, chodzi mi o końcówkę, "closer than 1", moze zamiast than 1 powiedziec, "than neutral non-human words"? albo jakoś podobnie} That such a move is important is suggested for instance by Figure 1 prepared using the original word list for religion-related protected words extended with control attributes, where such human attributes are also closer to protected classes than 1.

Another, also serious problem, is that  all the authors calculate means of means and their authors \todo{jak rozumiec drugi raz powtorzone their authors?} run statistical tests on sets of means. Unfortunately, by pre-averaging the data they throw away information about sample sizes, and  they remove variation, and so pre-averaging  tends to manufacture false confidence.

@@

\noindent
\begin{tabular}{ll}
\begin{minipage}[c]{0.35\linewidth}
\scalebox{0.65}{
\begin{tabular}{cccccc}\toprule
$X$ to $A$  & $X$ to $B$ & $Y$ to $A$ & $Y$ to $B$ & $\sigma$  & WEAT \\
\midrule
0 & 0 & 0 & 0 & 0.05 & 1.82 \\
0 & 0 & 0 & 0 & 0.001 & -1.93\\
.1 & 0 & 0 & .1 & 0.05 & 1.49 \\
.1 & 0 & 0 & .1 & .1 & 1.22\\
\bottomrule
\end{tabular}
}
\end{minipage}& \begin{minipage}[c]{0.6\linewidth}
\todo{początek zdania "two protected classes" bez wstepu typu "We created..", nie wiem czy tak miało być?}
Two protected classes, $X=\{t_1,t_2\}$ and $Y=\{t_3,t_4\}$, and two five-element attribute sets $A$ and $B$. One simple simulation draws normally distributed values for such combinations for two cases in which the underlying mean similarities are in fact equal, two cases in which they are different, and the underlying variances are different (code available upon request).
\end{minipage}
\end{tabular}

@@

The following observations are worthwhile. (1) For points randomly drawn from distributions where there is no difference in means the calculated effect size can easily be  1.82, whereas the largest effect size reported by @Caliskan2017semanticsBiases is 1.81. (2) For samples from distributions where the means are different, the (absolute) effect sizes can easily be lower than in the first two simulations.  As Figure 2. illustrates,  quite some uncertainty is involved, far from what systematically low mean-based p-values reported in the papers might suggest. Part of the problem is random variation unaccounted for in the original approach (see Figures 3 and 4 for an example), and part of the problem is that non-negligible changes in effect size can result from a shift in the standard deviation of the original process, because with the decrease of standard deviation the numerator in (WEAT) decreases.

@@

We build analogous models for other topic groups (race, gender), and different embeddings (Google, and hard-debiased Reddit). The general conclusion is that once we proceed this way, the situation is much less obvious. The word list sizes are small, sample sizes are small, and so posterior density intervals are wide. Some stronger bias can be discerned in the gender case. Moreover, sometimes the differences between associated, different and human, are not very impressive. The cost of debiasing turns out sometimes to be that   neutral human predicates get closer to protected classes (in religion),  change in gender after debiasing is really minor with   zero still out out of HPDI range, and debiasing improvement for race is a small improvement   achieved at the price of moving neutral terms closer to protected words (there is no space to visualise these results in this abstract).

### 2021-06-10 abstractESSLLI21/abstractESSLLI1.tex
Adds abstract review

\documentclass[12pt,dvipsnames,enabledeprecatedfontcommands]{scrartcl}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            colorlinks=true,
            linkcolor=Maroon,
            citecolor=Blue,
            urlcolor=blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

%\documentclass{article}

@@

%\bibliographystyle{apalike}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

@@

\begin{align}
\mathsf{s}(t,A,B)  = \frac{\sum_{a\in A}f(t,a)}{\vert A\vert} - \frac{\sum_{b\in B}f(t,b)}{\vert B\vert}
& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,
\mathsf{bias}(A,B)  = \frac{
\mu(\{s(x,A,B)\}_{x\in X}) -\mu(\{s(y,A,B)\}_{y\in Y}) 
}{
\sigma(\{s(w,A,B)\}_{w\in X\cup Y})
} \tag{WEAT}
\end{align}

\normalsize
\noindent {[}2{]} show that significant biases---thus measured---
similar to the ones discovered by IAT can be discovered in word
embeddings. {[}5{]} extends the methodology to a multilingual and
cross-lingual setting; a similar methodology is employed by {[}3{]}, who
employ word embeddings trained on corpora from different decades.
{[}6{]} modify WEAT to a multi-class setting, introducing Mean Average
Cosine (MAC) similarity as a measure of bias (in fact, in the paper they
report distances rather than similarities). Let
\(T = \{t_1, \dots, t_k\}\) be a class of protected word embeddings, and
let each \(A_j\in A\) be a set of attributes stereotypically associated
with a protected word). Then:

@@

\begin{align}
\mathsf{S}(t_i, A_j)  = \frac{1}{\vert A_j\vert}\sum_{a\in A_j}\mathsf{cos}(t,a) & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,
\mathsf{MAC}(T,A)  = \frac{1}{\vert T \vert \,\vert A\vert}\sum_{t_i \in T }\sum_{A_j \in A} S(t_i,A_j) \tag{MAC}
\end{align}

\normalsize  \noindent That is, for each protected word and each
attribute class, they first take the mean for this protected word and
all attributes in a given attribute class, and then take the mean of
thus obtained means for all the protected words.

Such methods are statistically problematic. One issue, specific to
{[}6{]}, is that no distinction is made based on whether given a
protected word and a class of attributes is stereotypically associated
with this protected word or a different protected word. A related,
general problem, is that all the authors ignore the step of comparing
their results with control groups, especially control groups of
stereotype-neutral human attributes. That such a move is important is
suggested for instance by Figure 1 prepared using the original word list
for religion-related protected words extended with control attributes,
where such human attributes are also closer to protected classes than 1.

Another, also serious problem, is that all the authors calculate means
of means and their authors run statistical tests on sets of means.
Unforutnately, by pre-averaging the data they throw away information
about sample sizes, and they remove variation, and so pre-averaging
tends to manufacture false confidence.

@@

\begin{tabular}{ll}
\begin{minipage}[c]{0.35\linewidth}
\scalebox{0.65}{
\begin{tabular}{cccccc}\toprule
$X$ to $A$  & $X$ to $B$ & $Y$ to $A$ & $Y$ to $B$ & $\sigma$  & WEAT \\
\midrule
0 & 0 & 0 & 0 & 0.05 & 1.82 \\
0 & 0 & 0 & 0 & 0.001 & -1.93\\
.1 & 0 & 0 & .1 & 0.05 & 1.49 \\
.1 & 0 & 0 & .1 & .1 & 1.22\\
\bottomrule
\end{tabular}
}
\end{minipage}& \begin{minipage}[c]{0.6\linewidth}
Two protected classes, $X=\{t_1,t_2\}$ and $Y=\{t_3,t_4\}$, and two five-element attribute sets $A$ and $B$. One simple simulation draws normally distributed values for such combinations for two cases in which the underlying mean similarities are in fact equal, two cases in which they are different, and the underlying variances are different (code available upon request).
\end{minipage}
\end{tabular}

@@

The following observations are worthwhile. (1) For points randomly drawn
from distributions where there is no difference in means the calculated
effect size can easily be 1.82, whereas the largest effect size reported
by {[}2{]} is 1.81. (2) For samples from distributions where the means
are different, the (absolute) effect sizes can easily be lower than in
the first two simulations. As Figure 2. illustrates, quite some
uncertainty is involved, far from what systematically low mean-based
p-values reported in the papers might suggest. Part of the problem is
random variation unaccounted for in the original approach (see Figures 3
and 4 for an example), and part of the problem is that that
non-negligible changes in effect size can result from a shift in the
standard deviation of the original process, because with the decrease of
standard deviation the numerator in (WEAT) decreases.

@@

\begin{align}
cosineDistance_i  & \sim dnorm(\mu_i, \sigma) \\
\mu_i & = m_{pw} + co_{con}\\
m_{pw} & \sim dnorm(1,.5) \\
co_{con} & \sim dnorm(0,1) \\
\sigma &\sim  dcauchy(0,1)
\end{align}

\normalsize

@@

\hypertarget{refs}{}
\hypertarget{ref-bolukbasi2016man}{}
{[}1{]} Tolga Bolukbasi, Kai-Wei Chang, James Zou, Venkatesh Saligrama,
and Adam Kalai. 2016. Man is to computer programmer as woman is to
homemaker? Debiasing word embeddings.

\hypertarget{ref-Caliskan2017semanticsBiases}{}
{[}2{]} Aylin Caliskan, Joanna J. Bryson, and Arvind Narayanan. 2017.
Semantics derived automatically from language corpora contain human-like
biases. \emph{Science} 356, 6334 (April 2017), 183--186.
DOI:\url{https://doi.org/10.1126/science.aal4230}

\hypertarget{ref-Garg2018years}{}
{[}3{]} Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James Zou.
2018. Word embeddings quantify 100 years of gender and ethnic
stereotypes. \emph{Proceedings of the National Academy of Sciences} 115,
16 (April 2018), E3635--E3644.
DOI:\url{https://doi.org/10.1073/pnas.1720347115}

\hypertarget{ref-Gonen2019lipstick}{}
{[}4{]} Hila Gonen and Yoav Goldberg. 2019. Lipstick on a pig: Debiasing
methods cover up systematic gender biases in word embeddings but do not
remove them. In \emph{Proceedings of the 2019 conference of the north
American chapter of the association for computational linguistics: Human
language technologies, volume 1 (long and short papers)}, Association
for Computational Linguistics, Minneapolis, Minnesota, 609--614.
DOI:\url{https://doi.org/10.18653/v1/N19-1061}

\hypertarget{ref-Lauscher2019multidimensional}{}
{[}5{]} Anne Lauscher and Goran Glavas. 2019. Are we consistently
biased? Multidimensional analysis of biases in distributional word
vectors. \emph{CoRR} abs/1904.11783, (2019). Retrieved from
\url{http://arxiv.org/abs/1904.11783}

\hypertarget{ref-Manzini2019blackToCriminal}{}
{[}6{]} Thomas Manzini, Yao Chong Lim, Yulia Tsvetkov, and Alan W Black.
2019. Black is to criminal as caucasian is to police: Detecting and
removing multiclass bias in word embeddings.

\hypertarget{ref-Nosek2002harvesting}{}
{[}7{]} Brian A. Nosek, Mahzarin R. Banaji, and Anthony G. Greenwald.
2002. Harvesting implicit group attitudes and beliefs from a
demonstration web site. \emph{Group Dynamics: Theory, Research, and
Practice} 6, 1 (2002), 101--115.
DOI:\url{https://doi.org/10.1037/1089-2699.6.1.101}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  12pt,
  dvipsnames,enabledeprecatedfontcommands]{scrartcl}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  colorlinks=true,
  linkcolor=Maroon,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=blue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
%\documentclass{article}

@@

%\bibliographystyle{apalike}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

\todo{czy konieczne jest pisanie "number vectors"? jeżeli tak, to w licencjacie musze pozmieniac, bo tam czasami jest samo vectors, a czasami "they are assigned numerical values"}

@@

\footnotesize \begin{align}
\mathsf{s}(t,A,B)  = \frac{\sum_{a\in A}f(t,a)}{\vert A\vert} - \frac{\sum_{b\in B}f(t,b)}{\vert B\vert}
& \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,
\mathsf{bias}(A,B)  = \frac{
\mu(\{s(x,A,B)\}_{x\in X}) -\mu(\{s(y,A,B)\}_{y\in Y}) 
}{
\sigma(\{s(w,A,B)\}_{w\in X\cup Y})
} \tag{WEAT}
\end{align} \normalsize \noindent {[}2{]} show that significant
biases---thus measured--- similar to the ones discovered by IAT can be
discovered in word embeddings. {[}5{]} extends the methodology to a
multilingual and cross-lingual setting; a similar methodology is
employed by {[}3{]}, who employ word embeddings trained on corpora from
different decades. {[}6{]} modify WEAT to a multi-class setting,
introducing Mean Average Cosine (MAC) similarity as a measure of bias
(in fact, in the paper they report distances rather than similarities).
Let \(T = \{t_1, \dots, t_k\}\) be a class of protected word embeddings,
and let each \(A_j\in A\) be a set of attributes stereotypically
associated with a protected word). Then:

@@

\begin{align}
\mathsf{S}(t_i, A_j)  = \frac{1}{\vert A_j\vert}\sum_{a\in A_j}\mathsf{cos}(t,a) & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,
\mathsf{MAC}(T,A)  = \frac{1}{\vert T \vert \,\vert A\vert}\sum_{t_i \in T }\sum_{A_j \in A} S(t_i,A_j) \tag{MAC}
\end{align} \normalsize  \noindent That is, for each protected word and
each attribute class, they first take the mean for this protected word
and all attributes in a given attribute class, and then take the mean of
thus obtained means for all the protected words.

Such methods are statistically problematic. One issue, specific to
{[}6{]}, is that no distinction is made based on whether given a
protected word and a class of attributes is stereotypically associated
with this protected word or a different protected word. A related,
general problem, is that all the authors ignore the step of comparing
their results with control groups, especially control groups of
stereotype-neutral human attributes.
\todo{nie wiem, czy nie przeredagowałabym tego zdania, chodzi mi o końcówkę, "closer than 1", moze zamiast than 1 powiedziec, "than neutral non-human words"? albo jakoś podobnie}
That such a move is important is suggested for instance by Figure 1
prepared using the original word list for religion-related protected
words extended with control attributes, where such human attributes are
also closer to protected classes than 1.

Another, also serious problem, is that all the authors calculate means
of means and their authors
\todo{jak rozumiec drugi raz powtorzone their authors?} run statistical
tests on sets of means. Unfortunately, by pre-averaging the data they
throw away information about sample sizes, and they remove variation,
and so pre-averaging tends to manufacture false confidence.

@@

\begin{tabular}{ll}
\begin{minipage}[c]{0.35\linewidth}
\scalebox{0.65}{
\begin{tabular}{cccccc}\toprule
$X$ to $A$  & $X$ to $B$ & $Y$ to $A$ & $Y$ to $B$ & $\sigma$  & WEAT \\
\midrule
0 & 0 & 0 & 0 & 0.05 & 1.82 \\
0 & 0 & 0 & 0 & 0.001 & -1.93\\
.1 & 0 & 0 & .1 & 0.05 & 1.49 \\
.1 & 0 & 0 & .1 & .1 & 1.22\\
\bottomrule
\end{tabular}
}
\end{minipage}& \begin{minipage}[c]{0.6\linewidth}
\todo{początek zdania "two protected classes" bez wstepu typu "We created..", nie wiem czy tak miało być?}
Two protected classes, $X=\{t_1,t_2\}$ and $Y=\{t_3,t_4\}$, and two five-element attribute sets $A$ and $B$. One simple simulation draws normally distributed values for such combinations for two cases in which the underlying mean similarities are in fact equal, two cases in which they are different, and the underlying variances are different (code available upon request).
\end{minipage}
\end{tabular}

@@

The following observations are worthwhile. (1) For points randomly drawn
from distributions where there is no difference in means the calculated
effect size can easily be 1.82, whereas the largest effect size reported
by {[}2{]} is 1.81. (2) For samples from distributions where the means
are different, the (absolute) effect sizes can easily be lower than in
the first two simulations. As Figure 2. illustrates, quite some
uncertainty is involved, far from what systematically low mean-based
p-values reported in the papers might suggest. Part of the problem is
random variation unaccounted for in the original approach (see Figures 3
and 4 for an example), and part of the problem is that non-negligible
changes in effect size can result from a shift in the standard deviation
of the original process, because with the decrease of standard deviation
the numerator in (WEAT) decreases.

@@

\begin{align}
cosineDistance_i  & \sim dnorm(\mu_i, \sigma) \\
\mu_i & = m_{pw} + co_{con}\\
m_{pw} & \sim dnorm(1,.5) \\
co_{con} & \sim dnorm(0,1) \\
\sigma &\sim  dcauchy(0,1)
\end{align} \normalsize

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\hypertarget{ref-bolukbasi2016man}{}%
\CSLLeftMargin{{[}1{]} }
\CSLRightInline{Tolga Bolukbasi, Kai-Wei Chang, James Zou, Venkatesh
Saligrama, and Adam Kalai. 2016. Man is to computer programmer as woman
is to homemaker? Debiasing word embeddings. Retrieved from
\url{http://arxiv.org/abs/1607.06520}}

\leavevmode\hypertarget{ref-Caliskan2017semanticsBiases}{}%
\CSLLeftMargin{{[}2{]} }
\CSLRightInline{Aylin Caliskan, Joanna J. Bryson, and Arvind Narayanan.
2017. Semantics derived automatically from language corpora contain
human-like biases. \emph{Science} 356, 6334 (April 2017), 183--186.
DOI:https://doi.org/\href{https://doi.org/10.1126/science.aal4230}{10.1126/science.aal4230}}

\leavevmode\hypertarget{ref-Garg2018years}{}%
\CSLLeftMargin{{[}3{]} }
\CSLRightInline{Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James
Zou. 2018. Word embeddings quantify 100 years of gender and ethnic
stereotypes. \emph{Proceedings of the National Academy of Sciences} 115,
16 (April 2018), E3635--E3644.
DOI:https://doi.org/\href{https://doi.org/10.1073/pnas.1720347115}{10.1073/pnas.1720347115}}

\leavevmode\hypertarget{ref-Gonen2019lipstick}{}%
\CSLLeftMargin{{[}4{]} }
\CSLRightInline{Hila Gonen and Yoav Goldberg. 2019. Lipstick on a pig:
{D}ebiasing methods cover up systematic gender biases in word embeddings
but do not remove them. In \emph{Proceedings of the 2019 conference of
the north {A}merican chapter of the association for computational
linguistics: Human language technologies, volume 1 (long and short
papers)}, Association for Computational Linguistics, Minneapolis,
Minnesota, 609--614.
DOI:https://doi.org/\href{https://doi.org/10.18653/v1/N19-1061}{10.18653/v1/N19-1061}}

\leavevmode\hypertarget{ref-Lauscher2019multidimensional}{}%
\CSLLeftMargin{{[}5{]} }
\CSLRightInline{Anne Lauscher and Goran Glavas. 2019. Are we
consistently biased? Multidimensional analysis of biases in
distributional word vectors. \emph{CoRR} abs/1904.11783, (2019).
Retrieved from \url{http://arxiv.org/abs/1904.11783}}

\leavevmode\hypertarget{ref-Manzini2019blackToCriminal}{}%
\CSLLeftMargin{{[}6{]} }
\CSLRightInline{Thomas Manzini, Yao Chong Lim, Yulia Tsvetkov, and Alan
W Black. 2019. Black is to criminal as caucasian is to police: Detecting
and removing multiclass bias in word embeddings. Retrieved from
\url{http://arxiv.org/abs/1904.04047}}

\leavevmode\hypertarget{ref-Nosek2002harvesting}{}%
\CSLLeftMargin{{[}7{]} }
\CSLRightInline{Brian A. Nosek, Mahzarin R. Banaji, and Anthony G.
Greenwald. 2002. Harvesting implicit group attitudes and beliefs from a
demonstration web site. \emph{Group Dynamics: Theory, Research, and
Practice} 6, 1 (2002), 101--115.
DOI:https://doi.org/\href{https://doi.org/10.1037/1089-2699.6.1.101}{10.1037/1089-2699.6.1.101}}

\end{CSLReferences}
